---
title: "Data_Science_Project"
author: "Juliana L. M"
date: "2025-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=FALSE}
getwd()
```
```{r}
library(tidyverse)
library(dplyr)
library(dslabs)
```

```{r}
maternal <- read.csv(file = 'Maternal Health Risk Data Set.csv')
head(maternal)
```
1.	Group the data by risk level 
(i)	Find the mean age of each risk level and order them in descending order.

```{r}
maternal %>% 
  group_by(RiskLevel) %>% summarize(avg = mean(Age)) %>% arrange(desc(avg))
```

```{r}
maternal %>% 
  arrange(desc(Age)) %>% 
  head()
```
1.	Group the data by risk level 
(ii)	Find the median diastolic of each risk level and order them in ascending order. 
```{r}
maternal %>% 
  group_by(RiskLevel) %>% summarize(DBpMedian = median(DiastolicBP)) %>% arrange(DBpMedian)
```
2.	Create a new variable name it systolic rate. Where systolic rate = systolic BP/age
```{r}
maternal <- mutate(maternal, systolic_rate = SystolicBP / Age)
head(maternal)
```
3.	Filter participants of age <= 40 years, determine average systolic rate for each risk level
```{r}
filter(maternal, Age <= 40) 
```
```{r}
filter(maternal, Age <= 40) %>% group_by(RiskLevel) %>% summarize(avgSR = mean(systolic_rate))
```
4.	For each continuous variable in the dataset, create histograms and box plots, and provide insights on the data distribution.
```{r}
hist(maternal$Age)
hist(maternal$DiastolicBP)
hist(maternal$SystolicBP)
hist(maternal$systolic_rate)
hist(maternal$BS)
hist(maternal$BodyTemp)
hist(maternal$HeartRate)
```
```{r}
boxplot(Age~RiskLevel, data = maternal)
boxplot(DiastolicBP~RiskLevel, data = maternal)
boxplot(SystolicBP~RiskLevel, data = maternal)
boxplot(systolic_rate~RiskLevel, data = maternal)
boxplot(BS~RiskLevel, data = maternal)
boxplot(BodyTemp~RiskLevel, data = maternal)
boxplot(HeartRate~RiskLevel, data = maternal)
```
5.	A) Explore the relationship between heart rate and the following variables using suitable graphical methods:
(i) Age
(ii) Systolic blood pressure
(iii) Diastolic blood pressure
(iv) Body temperature

```{r}
plot(maternal$Age, maternal$HeartRate) # Observed there is no linear relationship between Age and HeartRate
plot(maternal$SystolicBP, maternal$HeartRate)
plot(maternal$DiastolicBP, maternal$HeartRate)
plot(maternal$BodyTemp, maternal$HeartRate)
```
Calculate the five-number summary for each variable, grouped by risk level.
```{r}
summary(maternal)
```
B) Fit a simple linear model using heart rate as a response variable and the variables in A as predictor variables.
```{r}
modelAge <- lm (HeartRate~Age, data = maternal) #Responsive/Dependent variable vs Explanatory/Independent variable
modelSBP <- lm (HeartRate~SystolicBP, data = maternal)
modelDBP <- lm (HeartRate~DiastolicBP, data = maternal)
modelBTemp <- lm (HeartRate~BodyTemp, data = maternal)
summary(modelAge) #Display the model
summary(modelSBP)
summary(modelDBP)
summary(modelBTemp)

#It was observed that all models presented had the Adjusted R-Squared of less than 1% which is less than the recommended value of 60% and above. 
#Less than 1% of the variance in heart rate is explained by Age, SystolicBP, DiastolicBP and BodyTemp.
#Thus, the model is not fit to the data. 
```
6.	Challenge yourself: Fit a multiple linear regression model with heart rate as the response variable and the four predictor variables from Question 5. Then, apply stepwise selection to identify the best-fitting model.

# For all continuous variables against HeartRate
```{r}
model <- lm (HeartRate~Age + SystolicBP + DiastolicBP + BodyTemp, data = maternal)
summary(model)
#The model in comparison of all continuous variables against heart rate is particular significant since p-value: 2.51e-05 = 0.0000251 which is within the recommended range p-value < 0.05. 
#However,the model is overall not fit to these data since the Adjusted R-square is less than 60% 0.6 for all variables
```

# Age and BodyTemp against HeartRate
```{r}
modelBTMA <- lm (HeartRate~Age + BodyTemp, data = maternal)
summary(modelBTMA)
```


